Title
========================================================

Write up the assignment solution approach

A step wise approach was taken for this.

Step 1.Data Understanding  
In this step an attempt was made to understand the data.  This was achieved by reading the corresponding reference article and also examining the data points.  On analysis it was found there were many columns which had empty or NA values. There were also columns such as time stamps and user ids which will not add any value to the prediction in hand

Step 2.   Pre-Processing
In this step all columns which did not add any value to the prediction were removed from the data set.
Below is the list of all such columns
"timestamp|X|user_name|new_window|kurtosis_roll_belt|kurtosis_picth_belt|kurtosis_yaw_belt|skewness_roll_belt|skewness_roll_belt.1|skewness_yaw_belt | max_roll_belt | max_picth_belt | max_yaw_belt | min_roll_belt |	min_pitch_belt	| min_yaw_belt | amplitude_roll_belt |	amplitude_pitch_belt |	amplitude_yaw_belt | var_total_accel_belt | avg_roll_belt | stddev_roll_belt | var_roll_belt | avg_pitch_belt | stddev_pitch_belt | var_pitch_belt | avg_yaw_belt |	stddev_yaw_belt | var_yaw_belt | var_accel_arm |avg_roll_arm |	stddev_roll_arm | var_roll_arm | avg_pitch_arm |stddev_pitch_arm | var_pitch_arm | avg_yaw_arm | stddev_yaw_arm | var_yaw_arm | kurtosis_roll_arm | kurtosis_picth_arm | kurtosis_yaw_arm | skewness_roll_arm | skewness_pitch_arm | skewness_yaw_arm | max_roll_arm | max_picth_arm | max_yaw_arm | min_roll_arm | min_pitch_arm | min_yaw_arm | amplitude_roll_arm | amplitude_pitch_arm | amplitude_yaw_armn | kurtosis_roll_dumbbell | kurtosis_picth_dumbbell | kurtosis_yaw_dumbbell | skewness_roll_dumbbell | skewness_pitch_dumbbell | skewness_yaw_dumbbell | max_roll_dumbbell | max_picth_dumbbell | max_yaw_dumbbell | min_roll_dumbbell | min_pitch_dumbbell | min_yaw_dumbbell | amplitude_roll_dumbbell | amplitude_pitch_dumbbell	| amplitude_yaw_dumbbell | total_accel_dumbbell | var_accel_dumbbell| avg_roll_dumbbell | stddev_roll_dumbbell | var_roll_dumbbell | avg_pitch_dumbbell | stddev_pitch_dumbbell | var_pitch_dumbbell | avg_yaw_dumbbell | stddev_yaw_dumbbell | var_yaw_dumbbell | gyros_dumbbell_x | kurtosis_roll_forearm | kurtosis_picth_forearm | kurtosis_yaw_forearm	| skewness_roll_forearm	| skewness_pitch_forearm| skewness_yaw_forearm | max_roll_forearm | max_picth_forearm | max_yaw_forearm | min_roll_forearm | min_pitch_forearm | min_yaw_forearm | amplitude_roll_forearm | amplitude_pitch_forearm | amplitude_yaw_forearm  | var_accel_forearm | avg_roll_forearm | stddev_roll_forearm | var_roll_forearm | avg_pitch_forearm | stddev_pitch_forearm | var_pitch_forearm | avg_yaw_forearm | stddev_yaw_forearm | var_yaw_forearm"


Data was also split into training and validation sets.

Below is the pre-processing code snippet.
```{r}
library(caret)
set.seed(42)
# load data
trainRawData <- read.csv("pml-training.csv",na.strings=c("NA",""))
# discard NAs
# discard NAs
NAs <- apply(trainRawData,2,function(x) {sum(is.na(x))}) 
validData <- trainRawData[,which(NAs == 0)]



# make training set
trainIndex <- createDataPartition(y = validData$classe, p=0.2,list=FALSE) # 3927 rows
trainData <- validData[trainIndex,]
# discards unuseful predictors
removeIndex <- grep("timestamp|X|user_name|new_window|kurtosis_roll_belt|kurtosis_picth_belt|kurtosis_yaw_belt|skewness_roll_belt|skewness_roll_belt.1|skewness_yaw_belt | max_roll_belt | max_picth_belt | max_yaw_belt | min_roll_belt |  min_pitch_belt	| min_yaw_belt | amplitude_roll_belt |	amplitude_pitch_belt |	amplitude_yaw_belt | var_total_accel_belt | avg_roll_belt | stddev_roll_belt | var_roll_belt | avg_pitch_belt | stddev_pitch_belt | var_pitch_belt | avg_yaw_belt |	stddev_yaw_belt | var_yaw_belt | var_accel_arm |avg_roll_arm |	stddev_roll_arm | var_roll_arm | avg_pitch_arm |stddev_pitch_arm | var_pitch_arm | avg_yaw_arm | stddev_yaw_arm | var_yaw_arm | kurtosis_roll_arm | kurtosis_picth_arm | kurtosis_yaw_arm | skewness_roll_arm | skewness_pitch_arm | skewness_yaw_arm | max_roll_arm | max_picth_arm | max_yaw_arm | min_roll_arm | min_pitch_arm | min_yaw_arm | amplitude_roll_arm | amplitude_pitch_arm | amplitude_yaw_armn | kurtosis_roll_dumbbell | kurtosis_picth_dumbbell | kurtosis_yaw_dumbbell | skewness_roll_dumbbell | skewness_pitch_dumbbell | skewness_yaw_dumbbell | max_roll_dumbbell | max_picth_dumbbell | max_yaw_dumbbell | min_roll_dumbbell | min_pitch_dumbbell | min_yaw_dumbbell | amplitude_roll_dumbbell | amplitude_pitch_dumbbell	| amplitude_yaw_dumbbell | total_accel_dumbbell | var_accel_dumbbell| avg_roll_dumbbell | stddev_roll_dumbbell | var_roll_dumbbell | avg_pitch_dumbbell | stddev_pitch_dumbbell | var_pitch_dumbbell | avg_yaw_dumbbell | stddev_yaw_dumbbell | var_yaw_dumbbell | gyros_dumbbell_x | kurtosis_roll_forearm | kurtosis_picth_forearm | kurtosis_yaw_forearm	| skewness_roll_forearm	| skewness_pitch_forearm| skewness_yaw_forearm | max_roll_forearm | max_picth_forearm | max_yaw_forearm | min_roll_forearm | min_pitch_forearm | min_yaw_forearm | amplitude_roll_forearm | amplitude_pitch_forearm | amplitude_yaw_forearm  | var_accel_forearm | avg_roll_forearm | stddev_roll_forearm | var_roll_forearm | avg_pitch_forearm | stddev_pitch_forearm | var_pitch_forearm | avg_yaw_forearm | stddev_yaw_forearm | var_yaw_forearm",names(trainData))
trainData <- trainData[,-removeIndex]

print("removed non needed columns")
```

Step 3.  Fit the model.
In this step the model is fitted based on the training data. 

The random forest method has been used to fit the model due to it's high accuracy. 
The classe variable is the field to be predicted.

```{r}
modFit <- train(trainData$classe ~.,data = trainData,method="rf")

```


Step 4. Validation of model
In this step a prediction will be done using the model on the held out test set. The predicted output is compared with the actual value using the confusionMatrix.  The output was giving a satisfactory accuracy of
98.6% and Kappa score of .983.

```{r}
print("created the model")
testData <- validData[-trainIndex,]
testData <- testData[,-removeIndex]

# Predict with testdata
predictValues <- predict(modFit,newdata=testData)

#  Check the accuracy
confusionMatrix(testData$classe, predictValues)
```

Step 5.   Use the model for actual prediction
The test data would be loaded and will be pre-processed in the same way as training data. 
The fitted model will be used for predicting on this data set.

